---
title: "Day 5 Homework"
output: html_document
---

0. Reload your data, and re-use the code you already have for getting things formatted nicely (columns renamed, factors relabeled, etc.)

```{r}

rm(list=ls(all=TRUE))
library(readxl)
library(dplyr)
library(reshape2)

my.event.data<-read_excel("/Users/rithvikyarlagadda/Documents/R_GVPT622/PSYC798/my_R_homework/day3/event_data.xlsx")[,c("year","type_of_violence","conflict_name","dyad_name","where_coordinates", "country","best","adm_1")] #subsetting those columns from the main dataset that is required for my project

head(my.event.data)
summary(my.event.data)

#relabelling column names 
names(my.event.data)[names(my.event.data) == "type_of_violence"]<-"violence.type"
names(my.event.data)[names(my.event.data) == "conflict_name"]<-"conflict.name"
names(my.event.data)[names(my.event.data) == "dyad_name"]<-"dyad.name"
names(my.event.data)[names(my.event.data) == "where_coordinates"]<-"location.coordinates"
names(my.event.data)[names(my.event.data) == "best"]<-"fatalities"
names(my.event.data)[names(my.event.data) == "adm_1"]<-"province"

my.event.data$violence.type<-as.factor(my.event.data$violence.type) #this should actually be a factor variable but it is numeric. So, we convert into a factor 

my.event.data$violence.type<-factor(my.event.data$violence.type, levels = c(1,2,3), labels = c("state-based conflict", "non-state conflict", "one-sided violence")) #relabelling the names of the factor levels

levels(my.event.data$violence.type) #rechecking the levels of the factor variable 

```

1. Aggregate (e.g., get mean/median/max/something) of some numerical variable, over one or more factors.  Use `group_by()` plus `summarize()` from the `dplyr` package. You may or may not need to reshape the data first!

```{r}

#summarizing the fatalities variable (numeric) based on 3 different levels of violence type (factor) and different countries

violence_type.summaries <- summarize(group_by(my.event.data, violence.type, country), count.fatalities = sum(fatalities, na.rm=TRUE), mean.fatalities = mean(fatalities, na.rm = TRUE), median.fatalities = median(fatalities, na.rm = TRUE), maximum.fatalities = max(fatalities, na.rm = TRUE)) 

```

2. Get your data into "long" (aka "tidy") format, if it isn't already

```{r}

# The dataframe (my.event.data) is already in long (aka "tidy") format 

```







